---
title: "Code_TRF_Survival"
author: "Jorge Martinez Romero"
date: "`r format(Sys.time(), '%Y %B %d')`"
output: html_document
---
date: "`r format(Sys.time(), '%Y %B %d')`"


## Aim

Compare 5 interventions in mice, same strain and sex.

## Statistical Analysis 

Steps
1. Check mortality discrepancies in the dataframe (Memy)
2. Censor sacrificed animals 
3. Lifespan statistics (Without censored animals)
4. Survival Analysis: Chi squared and Log-rank test
5. Survival Analysis: Cox 
	 5.1 Check Assumptions
	 5.2 Fit parsimonious model
6. Compare survival in groups (Multiple hypothesis testing correction) 
	 6.1 Plots
7. Predict life expectancy according to the model.

## Results

***

```{r packages, warning=FALSE, message=FALSE, echo=FALSE}

library (survival)
library (kableExtra)
library (tidyverse)
library (survminer) 

knitr::opts_chunk$set(autodep = TRUE) 
knitr::opts_chunk$set(warning = FALSE) 
knitr::opts_chunk$set(message = FALSE) 
knitr::opts_chunk$set(echo = FALSE) 
knitr::opts_chunk$set(include = FALSE) 
knitr::opts_chunk$set(cache = FALSE) 

```


### Outcome: Lifespan 
<!-- LE -->
```{r life_expectancy, warning=FALSE, message=FALSE, echo=FALSE, include=TRUE}

#Import file cleaned and supplied by Memi
df<-read.csv2('Tables/TRF project_surv_stats.csv',sep = ";")
dfM<-read.csv2('Tables/TRF_fe_stats.csv',sep = ",")
dff<-df %>% select(idno, fu_age_wkCORRECTED,StatusCORRECTED)
dff2<-dfM %>% left_join(dff, by="idno") %>% 
  relocate(c(fu_age_wkCORRECTED,StatusCORRECTED),.after = fu_age_wk) %>% filter(feeding!="excluded")

dff2$feeding<-as.factor(as.character(dff2$feeding))


plot(density(df$fu_age_wkCORRECTED))
```


<!-- results -->
#### Normality
 
```{r shapi, warning=FALSE, message=FALSE, echo=FALSE, include=TRUE}
shapiro.test(df$fu_age_wkCORRECTED)
```

```{r lif, warning=FALSE, message=FALSE, echo=FALSE, include=TRUE}
library(ggpubr)
ggqqplot(df$fu_age_wkCORRECTED)
cat("No Normal, apply non parametric test or semi-parametric Cox for survival analysis")
```

## Lifespan Statistics (Sacrificed mice excluded)

```{r slam_ss_dist_funct, include=TRUE, warning=FALSE, message=FALSE, echo=FALSE, cache=TRUE, eval=TRUE}
#remove non natural death
df_nat<-df[df$StatusCORRECTED=="1",]

# From: Fernandez, Maria (NIH/NIA/IRP) [F] <fernandezme@mail.nih.gov> 
# Sent: Thursday, April 28, 2022 10:23 AM
# To: Martinez Romero, Jorge (NIH/NIA/IRP) [F] <jorge.martinezromero@nih.gov>
# Subject: RE: TRF4 y TRF8
# 
# Columna 1: feeding treatment (discriminando todos, i.e., AL, TRF4, TRF8, CR, CRX2).
# Columna 2: el ID de los animales.
# Columna 3: solo el numero de los IDs de los animales.
# Columna 4: la caja a la que pertenece cada animal (reconstruida a partir de uno de los archivos de Surv).
# Columna 5: la subcaja a la que pertenece cada animal (Tuve que hacer esto porque habia una caja “56A”).
# Columna 6: los animales que murieron antes de que empezaran los feeding treatments (estado=1) y los que murieron luego de que los feeding treatments habian comenzado (estado=0).
#Memi

#Bellow the file with diet assignation sent by Maria
df_5_diets<-read.csv2('Tables/Copy of TRFproject_IDAnimal_3 treatments.csv',sep = ",")
colnames(df_5_diets)[1]<-"Diet"
colnames(df_5_diets)[4]<-"Cage"

df_nat_5<-df_nat %>% left_join(df_5_diets, by = c("idno" = "ID")) %>% relocate(c(Diet,Cage),.after = feeding) 
df_nat_5$Diet[74:79]<-c(rep("CR",3),rep("CRX2",3))

df5<-df %>% left_join(df_5_diets, by = c("idno" = "ID")) %>% relocate(c(Diet,Cage),.after = feeding) 

df5$Diet[91:97]<-c(rep("CR",4),rep("CRX2",4))

Sumar<-as.matrix(aggregate(fu_age_wkCORRECTED ~ Diet, data = df_nat_5, summary))
#colnames(Sumar)
nn<-aggregate(fu_age_wkCORRECTED ~ Diet, data = df_nat_5, length)


Q90<-function(x){quantile(x, probs = 0.90, na.rm = T)}
QX90<-aggregate(fu_age_wkCORRECTED ~ Diet, data = df_nat_5, Q90)
colnames(nn)<-c("Diet","#Mice")
colnames(Sumar)[2:7]<-c("Min","Q1_25%","Median","Mean","Q3_75%","Max")
names(QX90)[2]<-"Q_90%"

Stats<-cbind(nn,Sumar[,2:6],QX90[2],Sumar[,7])
colnames(Stats)[9]<-"Max"
asNumeric <- function(x) as.numeric(as.character(x))
Stats<-Stats %>% mutate_if(is.factor, asNumeric)%>% mutate_if(is.numeric, round, digits=0)
Stats[,1]<-nn[,1]
#print(Stats)
kable(Stats, caption = "Table: Lifespan Statistics", row.names = FALSE, align = "c") %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE, position = "left") %>%
  # column_spec(8, border_right = T) %>%
  footnote(symbol = c("TRF:Time Restricted Feeding; CR: Caloric restriction; AL: Ad libitum")
           )
```

## Survival Analysis

### Chi squared and Log-rank test

```{r AA3, warning=FALSE, message=FALSE, echo=FALSE, include=TRUE} 

survdiff(Surv(time = fu_age_wkCORRECTED, StatusCORRECTED)~Diet,df5)

```
Conclusion:
The Chi squared test identifies sig. differences in diets, p<0.05


### Log-rank test
#### To check pairwise comparisons (with multiple testing adjustment)

#### TRF segregated in TRF4 TRF8
```{r AA4, warning=FALSE, message=FALSE, echo=FALSE, include=TRUE}
pairwise_survdiff(Surv(time = fu_age_wkCORRECTED, StatusCORRECTED)~Diet,df5, p.adjust.method = "BH", rho = 0)
```
No relevant results, to low n to perform this test

#### TRF merged together
```{r AA5, warning=FALSE, message=FALSE, echo=FALSE, include=TRUE}
pairwise_survdiff(Surv(time = fu_age_wkCORRECTED, StatusCORRECTED)~feeding,df5, p.adjust.method = "BH", rho = 0)

```
Same.

Conclusion:
In this case not enough Sample to identify differences by KM-Log rank pairwise analysis.


### Cox Proportional Hazard analysis

#### 1) Compare model TRF segregated vs TRF merged
#### 2) Find the parsimonious Cox model and check assumptions
```{r Modeling, echo=FALSE, eval=TRUE, include=TRUE, warning=FALSE, include=TRUE}
#Coxph with time restricted feeding split in 4 hrs TRF4 window and 8 hr window TRF8
fit_hr<-coxph(Surv(time = fu_age_wkCORRECTED, StatusCORRECTED)~Diet,method = "breslow",df5)
#Coxph with time restricted feeding grouped as TRF
fit_hr_2<-coxph(Surv(time = fu_age_wkCORRECTED, StatusCORRECTED)~feeding,method = "breslow",df5)
#Compare the two models
bb<-anova(fit_hr,fit_hr_2)
#They are similar pval=0.386. Keep parsimonious one TRF grouped
print(bb)
```

When comparing model with TRF clustered -> less loglik.
There is no difference between model TRF4+TRF8 and TRF. p=0.383 
No justification to split TRF. Keep the simplest model TRF clustered

#### Check proportional Hazard asumption
```{r Check proportional hazard ratios, echo=FALSE, eval=TRUE, include=TRUE, warning=FALSE, include=TRUE}
#Check proportional hazard assumption Schoefeld test
cox.zph(fit_hr_2)
plot(cox.zph(fit_hr_2)[1])
#Correct, they are proportional p>0.05 for model fit_hr_2. 
```
Risk seems to be proportional pva>0.05.
Plot strait. 

### Survival Comparison among diets (n=244, sacrificed censored)

#### Contrast all diet vs AL

```{r A1, warning=FALSE, message=FALSE, echo=FALSE, include=TRUE} 

summary(fit_hr_2)
```
Conclusion:
The Cox analysis finds one diet significantly altering survival risk, CR is protective compared with AL.
HR=0.55 (0.34-0.88) pval = 0.01*. The risk of dying under CR is less than a half compared to the risk undergoing AL.


### Contrast all diet vs CR
```{r A2, warning=FALSE, message=FALSE, echo=FALSE, include=TRUE} 
dfrl<-df5
dfrl$feeding<-factor(dfrl$feeding,levels=c('CR','CRX2','TRF','AL'))
fit_hr_3<-coxph(Surv(time = fu_age_wkCORRECTED, StatusCORRECTED)~feeding,dfrl)
summary(fit_hr_3)
```
Conclusion:
The Cox analysis finds 2 diets significantly altering survival risk, TRF increases risk compared with CR.
HR=1.75 (1.15-2.65) pval = 0.008 **. The risk of dying under TRF increases 75% (or 1.75 Fold) compared to the risk under CR.
AL increases risk compared with CR as seen in previous contrast * 


### Contrast all diet vs CRX2
```{r A3, warning=FALSE, message=FALSE, echo=FALSE, include=TRUE} 
dfrl<-df5
dfrl$feeding<-factor(dfrl$feeding,levels=c('CRX2','TRF','AL','CR'))
fit_hr_3<-coxph(Surv(time = fu_age_wkCORRECTED, StatusCORRECTED)~feeding,dfrl)
summary(fit_hr_3)
```
Conclusion:
No statistical significant results.

### Contrast all diet vs TRF
```{r A4, warning=FALSE, message=FALSE, echo=FALSE, include=TRUE} 
dfrl<-df5
dfrl$feeding<-factor(dfrl$feeding,levels=c('TRF','AL','CR','CRX2'))
fit_hr_3<-coxph(Surv(time = fu_age_wkCORRECTED, StatusCORRECTED)~feeding,dfrl)
summary(fit_hr_3)
```
Conclusion:
Confirmed what we saw in previous contrast.


### Plots
#### Kaplan-Meier curves model (TRF segregated)
```{r surv_curve5557, warning=FALSE, message=FALSE, echo=FALSE, include=TRUE}
plofit<-survfit(Surv(time = fu_age_wkCORRECTED, StatusCORRECTED) ~ Diet,df5)
library(survminer)
ggsurvplot(plofit, data = df5,
           legend = "top",
           legend.title = "Group: ",
           ylab = "Survivorship\n\n", 
           xlab = "\n\n Age (Weeks)", xlim = c(60,180),
           legend.labs = c("AL",
                           "CR","CRX2","TRF4","TRF8"),
           palette = c("Black","Dark Green","Blue","Dark Orange","red"),
           surv.median.line = "hv",
           ggtheme=theme_minimal())
```

### Kaplan-Meier curves  parsimonious model (TRF grouped)
```{r surv_curve5558, warning=FALSE, message=FALSE, echo=FALSE, include=TRUE}
plofit<-survfit(Surv(time = fu_age_wkCORRECTED, StatusCORRECTED) ~ feeding,df5)
library(survminer)
ggsurvplot(plofit, data = df5,
    legend.title = "Group: ",
           ylab = "Survivorship\n\n", 
           xlab = "\n\n Age (Weeks)", xlim = c(60,180),
           legend.labs = c("AL",
                           "CR","CRX2","TRF"),
           surv.median.line = "hv",
           palette = c("Black","Dark Green","Blue","DArk Orange"),
           ggtheme=theme_minimal())

```








